# Enumerate possible options {#args-enum}

```{r, include = FALSE}
source("common.R")
```

## What's the pattern?

If an argument can take one of a relatively small set of known values, specify all possible values in the default argument, and then use `match.arg()` or `rlang::arg_match()` to pick. This pattern is often naturally paired with a `switch()` to take action.

(See ???? for a similar pattern that can be extended by others.)

## What are some examples?

* `rank()` has a `ties.method` that describes the method used to handle the 
  rank of ties. It must be "average", "first", "last", "random", "max", or 
  "min". 

## Why is it important?

## How do I use it?

`rank()` documents this restriction with the by including a vector of possible values as the default.

```{r}
rank <- function(x, 
                 ties.method = 
                  c("average", "first", "last", "random", "max", "min")
                 ) {
  
  ties.method <- match.arg(ties.method)
  
  switch(ties.method, 
    average = , 
    min = , 
    max = .Internal(rank(x, length(x), ties.method)), 
    first = sort.list(sort.list(x)),
    last = sort.list(rev.default(sort.list(x, decreasing = TRUE))), 
    random = sort.list(order(x, stats::runif(length(x))))
  )
}

x <- c(1, 2, 2, 3, 3, 3)

rank(x)
rank(x, ties.method = "first")
rank(x, ties.method = "min")
```

Note that `match.arg()` will automatically throw an error if the value is not in the set:

```{r, error = TRUE}
rank(x, ties.method = "middle")
```

It also supports partial matching so that the following code is shorthand for `ties.method = "random":

```{r}
rank(x, ties.method = "r")
```

We generally believe that partial matching is a bad idea, because it makes code harder to read. `rlang::arg_match()` is an alternative to `match.args()` that doesn't support partial matching. Instead it provides a helpful error message:

```{r, error = TRUE}
rank2 <- function(x, 
                 ties.method = 
                  c("average", "first", "last", "random", "max", "min")
                 ) {
  
  ties.method <- rlang::arg_match(ties.method)
  rank(x, ties.method = ties.method)
}

rank2(x, ties.method = "r")
```

### How to use it and keep defaults short?

Best used when the set of possible values is short. You can see in the `rank()` example it's already getting unwieldy and makes it difficult to nicely format. The easiest way to follow Chapter \@ref(args-defaults-short) and extract the possible values into its own function:

```{r}
ties_method <- function() {
  c("average", "first", "last", "random", "max", "min")
}
```

(If in a package, this function should be exported and documented with the function that uses it.)

Then you can use that as a default argument:

```{r}
rank2 <- function(x, ties.method = ties_method()) {
  ...
}
```

This is not quite as good as inlining all possible values, since there's some indirection, so you can't see the possible values in tooltips, but you can still access programmatically.

This is also useful if you want to share the values across multiple functions. For example `stats::p.adjust()`, `stats::pairwise.prop.test()`, `stats::pairwise.t.test()`, `stats::pairwise.wilcox.test()` all use `p.adjust.method = p.adjust.methods`.
