# Avoid hidden arguments {#sec-args-hidden}

```{r}
#| include = FALSE
source("common.R")
```

## What's the problem?

Functions are easier to understand if the outputs depends only on the values of the input.
If a function unexpectedly returns different results with the same inputs, then we say it has **hidden arguments**.
Hidden arguments make code harder to reason about, because to correctly predict the output you also need to know some other state.

It's hard to precisely characterise surprise, but it's an important part of this pattern as their are functions whose primary purpose is to return different values even with the same inputs:

-   `Sys.time()` depends on the system time, but it's not a surprise: getting the current time is to the whole point of the function!

-   `read.csv(path)` depends not on the value of `path` but the contents of the file at that location.
    Reading from the file system necessarily implies that the results depend on the contents of the file, not its path, so this is not a surprise.

-   Random number generators like `runif()` peek at the value of the special global variable `.Random.seed`.
    This is a little surprising, but if they didn't have some global state every call to `runif()` would return the same value.

Related to hidden arguments (where a function depends on global state in a surprising way) is spooky action (@sec-spooky-action), where a function changes global state in a surprising way.

## What are some examples?

One common source of hidden arguments is the use of global options.
These can be useful to control display but, as discussed in @sec-def-user, should not affect computation:

-   The result of `data.frame(x = "a")$x` depends on the value of the global `stringsAsFactors` option: if it's `FALSE` (the default) you get a character vector; if it's `TRUE`, you get a factor.

-   `lm()`'s handling of missing values depends on the global option of `na.action`.
    The default is `na.omit` which drops the missing values prior to fitting the model (which is inconvenient because then the results of `predict()` don't line up with the input data.
    `modelr::na.warn()` provides an approach more in line with other base behaviours: it drops missing values with a warning.)

Another common source of hidden inputs is the system locale:

-   `strptime()` relies on the names of weekdays and months in the current locale.
    That means `strptime("1 Jan 2020", "%d %b %Y")` will work on computers with an English locale, and fail elsewhere.
    This is particularly troublesome for Europeans who frequently have colleagues who speak a different language.

-   `as.POSIXct()` depends on the current timezone.
    The following code returns different underlying times when run on different computers:

    ```{r}
    as.POSIXct("2020-01-01 09:00")
    ```

-   `toupper()` and `tolower()` depend on the current locale.
    It is fairly uncommon for this to cause problems because most languages either use their own character set, or use the same rules for capitalisation as English.
    However, this behaviour did cause a bug in ggplot2 because internally it takes `geom = "identity"` and turns it into `GeomIdentity` to find the object that actually does computation.
    In Turkish, however, the upper case version of i is İ, and `Geomİdentity` does not exist.
    This meant that for some time ggplot2 did not work on Turkish computers.

-   For similar reasons, `sort()` and `order()` rely on the lexicographic order defined by the current locale.
    `factor()` uses `order()`, so the results from factor depend implicitly on the current locale.
    (This is not an imaginary problem as this [SO question](https://stackoverflow.com/questions/39339489) attests).

## Why is it important?

Hidden arguments are bad because they make it much harder to predict the output of a function.
Historically, the worst offender was the `stringsAsFactors` option which changed how a number of functions[^args-hidden-1] treated character vectors.
This option was part of a multiyear procedure to move R away toward charcter vectors and away from vectors.
You can learn more in [*stringsAsFactors: An unauthorized biography*](https://simplystatistics.org/posts/2015-07-24-stringsasfactors-an-unauthorized-biography/) by Roger Peng and [*stringsAsFactors = \<sigh\>*](http://notstatschat.tumblr.com/post/124987394001/stringsasfactors-sigh) by Thomas Lumley.

[^args-hidden-1]: Such as `data.frame()`, `as.data.frame()`, and `read.csv()`

Allowing the system locale to affect the result of a function is a subtle source of bugs when sharing code between people who work in different countries.
These defaults on rarely cause problems because most languages that share the same writing system share (most of) the same collation rules.
The main exceptions tend to be European languages which have varying rules for modified letters, e.g. in Norwegian, å comes at the end of the alphabet.
However, when they do cause problems they will take a long time to track down: you're unlikely to expect that the coefficients of a linear model are different[^args-hidden-2] because your code is run in a different country!

[^args-hidden-2]: You'll get different coefficients for a categorical predictor if the ordering means that a different levels comes first in the alphabet.
    The predictions and other diagnostics won't be affected, but you're likely to be surprised that your coefficients are different.

## How can I remediate the problem?

Generally, hidden arguments are easy to avoid when creating new functions: simply refrain from using environment variables (like the locale) or global options (like `stringsAsFactors`).
The easiest way for problems to creep in is indirectly, when you call a function not knowing that it has hidden inputs.
The best way to prevent that is to consult the list of common offenders provided above.

If you want depend on an environment variable or option, make sure it's an explicit argument.
Such arguments generally should not affect computation (only side-effects like printed output or status messages); if they do affect results, follow @sec-def-inform to make sure the user knows what's happening.
For example, lets take `as.POSIXct()` which basically looks something like this:

```{r}
as.POSIXct <- function(x, tz = "") {
  base::as.POSIXct(x, tz = tz)
}
as.POSIXct("2020-01-01 09:00")
```

The `tz` argument is present, but it's not obvious that `""` means the currrent time zone.
Let's first make that explicit:

```{r}
as.POSIXct <- function(x, tz = Sys.timezone()) {
  base::as.POSIXct(x, tz = tz)
}
as.POSIXct("2020-01-01 09:00")
```

Since this is an important default whose value can change, we also print it out if the user hasn't explicitly set it:

```{r}
as.POSIXct <- function(x, tz = Sys.timezone()) {
  if (missing(tz)) {
    message("Using `tz = '", tz, "'`")
  }
  base::as.POSIXct(x, tz = tz)
}
as.POSIXct("2020-01-01 09:00")
```
