# Data, then details {#args-data-details}

```{r, include = FALSE}
source("common.R")
```

(If you're not familiar with the terms "data argument" and "details argument", read Chapter \@ref(call-data-details) before continuing.)

## What's the pattern?

When creating a function, put data arguments before details arguments. 

Arguments with defaults should always come after arguments without defaults (i.e. required arguments).

### Related patterns

It's possible for the data argument to be `...` (i.e. there's an arbitrary number of data arguments). For example, `paste()` allows you to supply any number of strings to `...`, and the details of the concatenation are controlled by `sep` and `collapse`. This pattern is best using sparingly, and is described in more detail in Chapter \@ref(dots-data). 

`...` can also be used to capture details arguments and pass them on to other functions. See Chapters \@ref(dots-position) and \@ref(dots-inspect) to how to use `...` as safely as possible in this situation.

Convention for calling Chapter \@ref(call-data-details).

## Why bother?

Works naturally with the pipe.

Convention makes it easier to understand structure of function at a glance - know that the most important arguments come first.

## What functions violate this pattern?

One larger family of functions in base R are the functions that tend not to follow this are the string manipulation functions. For example, `grepl(pattern, x)` returns a logical vector the same length as `x` containing `TRUE` wherever `pattern` is matched. Here, `x` is the data argument (it's the character vector of strings to compare), and `pattern` is the data argument. 

Another important example is `lm()`: I think it's fairly obvious that the primary data argument is `data` and one of the key details is the model specification in the `formula` argument. This is partly a historic accident, because having `lm()` work with data frames (rather than individual vectors in the current environment), did become important until later. 

`replicate()` takes the number of replicates as the first argument, whereas that seems more like a detail of the computation.

```{r}
replicate(3, rnorm(4))
replicate(rnorm(4), n = 3)
```

`setNames()` has a very unusual function:

```{r}
setNames <- function(object = nm, nm) {
  names(object) <- nm
  object
}
```

The purrr equivalent, `set_names()` adopts the more obvious specification:

```{r}
set_names <- function(x, nm = x) {
  names(x) <- nm
  x
}
```

(Here `x` is a better name than `object`, because only vectors can have names, and `x` is a convential argument name for a vector.)

(`set_names()` is actually a bit more complicated because the second argument can be a function used to transform existing names. See the documentation for more details.)

See also `mapply()` vs. `pmap()` as discussed in Chapter \@ref(cs-mapply-pmap).

## How do I recover?

If you have an exported function with this problem, there's no way to fix it. You'll need to deprecate the function and come up with a new replacement without the problem.
