# Hidden inputs {#args-hidden}

```{r, include = FALSE}
source("common.R")
```

## What's the problem?

Functions are easier to understand if the results depend only on the values of the inputs. If this is not true, and result of a function depends might be different with the same input arguments, it has __hidden inputs__.

#### Related patterns {-}

* This pattern is about surprising inputs; [Spooky action] is about
  suprising outputs.

## What are some examples?

* `data.frame()` sneakily peeks at the `stringsAsFactors` global option so 
  that the resulting types. If you look at `data.frame(x = letters)` in 
  isolation, you can't predict whether it the `x` variable will be a 
  factor or character vector.
  
* `lm()` + `na.action`.

Another common source of hidden inputs is the system locale:
  
* `strptime()` depends on the names of weekdays and months in the current 
  locale. That means `strptime("1 Jan 2020", "%d %b %Y")` will work on some
  computers and fail on others. This is particularly troublesome for Europeans
  who frequently work with other people who have different first language.
  
* `toupper()` and `tolower()` depend on the current locale. It is faily 
  uncommon for this to cause problems because most languages either
  use their own character set, or use the same rules for capitalisation as 
  English. However, this behaviour did cause a bug in ggplot2 because 
  internally it takes `geom = "identity"` and turns it into `GeomIdentity` 
  to find the object that actually does computation. In Turkish, however, the 
  upper case version of i is İ, and `Geomİdentity` does not exist.
  
* `sort()` and `order()` sneakily peak at `Sys.getlocale()`, and 
  `factor()` uses `order()`, so the results from factor depend implicitly 
  on the current locale.

Some functions depend on external settings, but not in a surprising way:
  
* `Sys.time()` depends on the system time, but it's not a surprise: getting
  the current time is to the whole point of the function!

* `read.csv(path)` depends not on the value of `path` but the contents of the 
  file at that location. Reading from the file system necessarily implies that
  the results depend on the contents of the file, not its path, so this is not
  a surprise.

* Random number generators like `runif()` peek at the value of the special 
  global variable `.Random.seed`. This is a little surprising, but if they 
  didn't have some global state every call to `runif()` would return the 
  same value.

Other functions depend on hidden inputs to control side-effects

* `print()` + `options(digits, width)` 

* `install.packages()` + `options(repos)`

This is fine - it gives the user control over appearance without affecting the computed values, and is discussed more in Chapter \@ref(args-user-defaults).

## Why is it bad?

Hidden inputs are bad because they make it much harder to predict the output of a fuction. The worst offender by far is the `stringsAsFactors` option which changes how a number of functions (including `data.frame()`, `as.data.frame()`, `expand.grid()`, and `read.csv()`) treat character vectors. This exists mostly for historical reasons, as described in recommend [*stringsAsFactors: An unauthorized biography*](http://simplystatistics.org/2015/07/24/stringsasfactors-an-unauthorized-biography/) by Roger Peng and [*stringsAsFactors = \<sigh\>*](http://notstatschat.tumblr.com/post/124987394001/stringsasfactors-sigh)
by Thomas Lumley. )

## How can I avoid it?

Avoiding hidden inputs is simple: ensure that the results of your function only depend on its input arguments.

If you need to read from a setting external to the function or to R itself, it should be an argument to the function, where the default value makes it clear that the value is coming from somewhere else (i.e. not `NULL` or not using `missing()`ness). This has the added benefit of also making it easier to test your code.

If you have many options that the user might want to vary, consider an explicit `locale` object like in readr.

You can also make the use of the input less hidden by explicitly notifying the user whenever they rely on the default value that might vary in surprising ways. `install.packges()` does this to notify you about it's use of the `.libPath()` option.

```{r, eval = FALSE}
install.packages("forcats")
# Installing package into ‘/Users/hadley/R’
# (as ‘lib’ is unspecified)
```

But it's easy to ignore this amongst the other output that `install.packages()` produces, and the message doesn't refer to the mechanism that controls it: `.libPaths()`.

## Related

* GITHUB_PAT
* auth and API keys
