# Avoid Boolean strategies {#sec-avoid-boolean-strategies}

```{r}
#| include = FALSE
source("common.R")
```

## What's the pattern?

If your function provides only two strategies it's tempting to distinguish between them using a Boolean argument that takes `TRUE` or `FALSE`.
I recommend against this unless you're REALLY SURE there won't ever be another strategy.
There are two reasons:

-   If you do discover a third strategy, you'll need to carefully remediate your function.
-   In many cases, you can be more express with a enumeration than a Boolean argument.

## What are some examples?

-   It's interesting to compare the interfaces of `sort()` and `vctrs::vec_sort()`:

    ```{r}
    #| results: false
    x <- sample(10)
    sort(x, decreasing = TRUE)
    vctrs::vec_sort(x, direction = "desc")

    sort(x, decreasing = FALSE)
    vctrs::vec_sort(x, direction = "asc")
    ```

    While it's unlikely that we'll ever need another sorting direction, I think `direction = "asc"`/`"desc"` is mildly easier to understand than `decreasing = FALSE`/`TRUE`.

-   The `na.last` argument to `sort()` adds `NA` to the mix to allow a Boolean argument to specify one of three possible strategies: put NAs at the end (`TRUE`), at the beginning (`FALSE`), or drop them alltogether (`NA`).

-   `haven::write_sav()` started with a compress argument that could be either `TRUE` (compress the file) or `FALSE` (don't compress the file). But then SPSS introduced a new way of compressing files leading to three possible options: compress with the new way, compress with the old way, or don't compress.

## How do you remediate past mistakes?

There are two possible cases, based on whether the old argument name still makes sense with the new argument values.
The sections below show what you'll need to do if you need a new argument (most cases) or if you can reuse the existing argument.

### New argument

Imagine we wanted to remediate the `na.last` argument to `sort()`:

-   `na.last = TRUE` means put `NA`s last.
-   `na.last = FALSE` means put `NA`s first.
-   `na.list = NA` means to drop them.

I think we could make this function more clear by instead changing the argument name to `na`, and then accepting one of three values: `last`, `first`, or `drop`.
To change an argument name in a backward compatible way, you deprecate the old argument and add a new argument, as below.

```{r}
sort <- function(x, 
                 na.last = lifecycle::deprecated(), 
                 na = c("drop", "first", "last")) 
{
  if (lifecycle::is_present(na.last)) {
    lifecycle::deprecate_warn("1.0.0", "sort(na.last)", "sort(na)")

    if (!is.logical(na.last) || length(na.last) != 1) {
      cli::cli_abort("{.arg na.last} must be a single TRUE, FALSE, or NA.")
    }
    
    if (isTRUE(na.last)) {
      na <- "last"
    } else if (isFALSE(na.last)) {
      na <- "first"
    } else {
      na <- "drop"
    }
  } else {
    na <- arg_match(na)
  }
  
  ...
}
```

::: callout-note
Note that because `na` is a prefix of `na.last` and `na.last` comes before `…` (in violation of @sec-dots-after-required), this introduces a very subtle behaviour change: a call like `sort(x, n = TRUE)` would have previously worked and been equivalent to `sort(x, na.last = TRUE)` but will now fail because it matches both `na` and `na.last`.

It would also be nice to make the default value `"last"` since it's very unusual for an R function to silently remove missing values.
However, that is much more likely to be a breaking change, affecting much existing code, so is unlikely to be worthwhile.
:::

### Re-use existing name

Sometimes you get lucky and you can use the same argument name, as with `haven::write_sav()`.
When SPSS version 21.0 introduced a new form of compression (called `zsav`), we change the compress argument from accepting `TRUE` or `FALSE` to accepting, "none", "byte", or "zsav" because `compress = TRUE` reads just as well as `compress = "byte"`.

To make this change in the code, we kept the existing behaviour for `TRUE` and `FALSE` while also adding the ability to supply a string:

```{r}
write_sav <- function(data, path, compress = c("byte", "none", "zsav"), adjust_tz = TRUE) {
  if (isTRUE(compress)) {
    compress <- "zsav"
  } else if (isFALSE(compress)) {
    compress <- "none"
  } else {
    compress <- arg_match(compress)
  }

  ...
}
```

You could imagine deprecating the old logical options, but here we chose to the keep them since it's a small amount of extra code, and means that existing users never need to worry about it.
See `?haven::read_sav` for how we communicated the in the docs.
