# Use a strategy, even for two choices {#sec-avoid-boolean-strategies}

```{r}
#| include = FALSE
source("common.R")
```

## What's the pattern?

If your function implements two strategies it's tempting to distinguish between them using an argument that takes either `TRUE` or `FALSE`.
I recommend against this unless you're **really sure** there won't ever be another strategy.
There are two reasons:

-   If you do discover a third strategy, you'll need to carefully remediate your function.
-   In many cases, you can be more express with a enumeration than a Boolean argument.

## What are some examples?

There are quite a few examples of the problem in tidyverse, because this is a pattern that we discovered relatively recently.
Here are two examples that I spotted recently:

-   `httr2::multi_req_perform()` has a very good example of this problem as it has an argument called `cancel_on_error`.
    In hindsight, it's very obvious that there are other things that you might want to do on error other than cancel, and even though I don't know what they are know, it would be better to design for the future.

-   `haven::write_sav()` started with a compress argument that could be either `TRUE` (compress the file) or `FALSE` (don't compress the file).
    But then SPSS introduced a new way of compressing files leading to three possible options: compress with the new way, compress with the old way, or don't compress.
    We'll discuss what we did to remediate the problem below.

Another interesting case is `sort()` which has two arguments that take a single logical value: `decreasing` and `na.last`.

-   `decreasing` is interesting to think about because it picks between two directions: ascending and descending.
    It's extremely unlikely that we'd ever need a third direction, but I think there's still benefit in using the strategy pattern here because I think it makes the arguments mildly easier to understand.
    For example, compare the interfaces of `sort()` and `vctrs::vec_sort()`:

    ```{r}
    #| results: false
    x <- sample(10)
    sort(x, decreasing = TRUE)
    vctrs::vec_sort(x, direction = "desc")

    sort(x, decreasing = FALSE)
    vctrs::vec_sort(x, direction = "asc")
    ```

    While you never need to explicitly right `decreasing = FALSE` (since that's the default), I think it's easier to understand `direction = "asc"` than trying to process what not-decreasing means.

-   The `na.last` argument to `sort()` can be `TRUE` (put `NA`s at the end), `FALSE` (put `NA`s at the beginning), or `NA` (drop `NA`s from the result).
    This is an interesting way to extend a Boolean argument to take one of three different values.

## How do you remediate past mistakes?

There are two possible ways to switch to using a strategy instead of `TRUE`/`FALSE` depending on whether the old argument name still makes sense with the new argument values.
The sections below show what you'll need to do if you need a new argument (most cases) or if you're lucky enough to be able to reuse the existing argument.

### New argument

Imagine we wanted to remediate the `na.last` argument to `sort()`:

-   `na.last = TRUE` means put `NA`s last.
-   `na.last = FALSE` means put `NA`s first.
-   `na.list = NA` means to drop them.

I think we could make this function more clear by changing the argument name to `na` and accepting one of three values: `last`, `first`, or `drop`.
Changing an argument name is equivalent removing the old name and adding the new name.
This way of thinking about the change makes it easier to see how you do it in a backward compatible way: you you just need deprecate the old argument in favour of the new one.

```{r}
sort <- function(x, 
                 na.last = lifecycle::deprecated(), 
                 na = c("drop", "first", "last")) 
{
  if (lifecycle::is_present(na.last)) {
    lifecycle::deprecate_warn("1.0.0", "sort(na.last)", "sort(na)")

    if (!is.logical(na.last) || length(na.last) != 1) {
      cli::cli_abort("{.arg na.last} must be a single TRUE, FALSE, or NA.")
    }
    
    if (isTRUE(na.last)) {
      na <- "last"
    } else if (isFALSE(na.last)) {
      na <- "first"
    } else {
      na <- "drop"
    }
  } else {
    na <- arg_match(na)
  }
  
  ...
}
```

::: callout-note
Note that because `na` is a prefix of `na.last` and `sort()` puts `na.last` before `…,`not after it (see @sec-dots-after-required), this introduces a very subtle behaviour change.
Previously, `sort(x, n = TRUE)` would have worked and been equivalent to `sort(x, na.last = TRUE)`.
But now it will fail, because `n` is a prefix of two arguments (`na` and `na.last)`.
This is unlikely to affect much code, but is something that you need to consider.

It would also be nice to make the default value `"last"` to match `order()`, especially since it's very unusual for a function to silently remove missing values.
However, that's likely to affect a lot of existing code, meaning that it's unlikely to be worthwhile.
:::

### Re-use existing name

When remediating a old mistake sometimes you get lucky and you can use the same argument name, as with `haven::write_sav()`.
When SPSS version 21.0 introduced a new form of compression (called `zsav`), we changed the compress argument from accepting `TRUE` or `FALSE` to accepting, "none", "byte", or "zsav" because `compress = TRUE` reads just as well as `compress = "byte"`.
To implement thiis we kept the existing behaviour for `TRUE` and `FALSE` while also adding the ability to supply a string:

```{r}
write_sav <- function(data, path, compress = c("byte", "none", "zsav"), adjust_tz = TRUE) {
  if (isTRUE(compress)) {
    compress <- "zsav"
  } else if (isFALSE(compress)) {
    compress <- "none"
  } else {
    compress <- arg_match(compress)
  }

  ...
}
```

You could choose to deprecate `TRUE` and `FALSE`, but here we chose to the keep them since it's only a small amount of extra code in haven, and it means that existing users don't need to think about it.
See `?haven::read_sav` for how we communicated the change in the docs.
