# Work with multiple dependency versions {#changes-multivers}

```{r, include = FALSE}
source("common.R")
```

## What's the pattern?

In an ideal world, when a dependency of your package changes its interface, you want your package to work with both versions. This is more work but it has two significant advantages:

* The CRAN submission process is decoupled. If your package only works with the
  development version of a dependency, you'll need to carefully coordinate your
  CRAN submission with the dependencies CRAN submission. If your package works
  with both versions, you can submit first, making life easier for CRAN and for
  the maintainer of the dependency.
  
* User code is less likely to be effected. If your package only works with the
  latest version of the dependency, then when a user upgrades your package, the
  dependency also must update. Upgrading multiple packages is more likely to
  affect user code than updating a single package.

In this pattern, you'll learn how to write code designed to work with multiple versions of a dependency, and you'll how to adapt your existing Travis configuration to test that you've got it right.

## Writing code

Sometimes there will be an easy way to change the code to work with both old and new versions of the package. However, in most cases, you'll need an `if` statement that uses one call for old versions of the dependency and one call for new versions of the package:

```{r, eval = FALSE}
if (dependency_has_new_interface()) {
  # freshly written code that works with new version
} else {
  # previously written code
}
```

Note that this code will usually generate an R CMD check `NOTE` when you submit it to CRAN because the fresly written code will use functions that don't exist in the CRAN version. This is one of the few NOTEs that you can explain to CRAN: just mention that it is needed for forward compatibility in your submission notes.

There are three basics approaches to test if a package has the new interface:

* Check the version of the package. This is recommended in most cases, 
  but requires that the dependency author use a specific version convention.
  
* Check for existence of a function.

* Check for existence or value of an argument.

### Case study: tidyr

To make the problem concrete so we can show of some real code, lets imagine we have a package that uses `tidyr::nest()`. Between versions 0.8.3 and 1.0.0, tidyr changed the interface of `nest()`. There are three versions of tidyr that we need to think about:

* 0.8.3: the version currently on CRAN with the old interface.

* 0.8.99.9000: the current development version with the new version. As 
  usually, the fourth component is >= 9000 to indicate that it's a development 
  version. Note, however, that the patch version is 99; this indicates that 
  release includes breaking changes.
  
* 1.0.0: the future CRAN version; this is the version that will be submitted 
  to CRAN.

And packages need to include code like this:

```{r, eval = FALSE}
if (tidyr_new_interface()) {
  out <- tidyr::nest_legacy(df, x, y, z)
} else {
  out <- tidyr::nest(df, c(x, y, z))
}
```

(As described above, when submitted to CRAN this will generate a note about missing `tidyr::nest_legacy()` which can be explained in the submission comments.)

The main question is how to write `tidyr_new_interface()`. There are three options:

*   Check that the version is greater than the development version:

    ```{r}
    tidyr_new_interface <- function() {
      packageVersion("tidyr") > "0.8.99"
    }
    ```
    
    This technique works because tidyr uses the convention that the 
    development version of backward incompatible functions contain `99`
    in the third (patch) component.

*   If tidyr didn't adopt this naming convention, we could test for the 
    existence of `unnest_legacy()`. 
  
    ```{r}
    tidyr_new_interface1 <- function() {
      exists("unnest_legacy", asNamespace("tidyr"))
    }
    ```
  
*   If the inteface change was more subtle, you might have to think more 
    creatively. If the package uses the [lifecycle](http://lifecycle.r-lib.org/)
    system, one approach would be to test for the presence of `deprecated()`
    in the function arguments:
    
    ```{r}
    tidyr_new_interface2 <- function() {
      identical(formals(tidyr::nest)$.key, quote(deprecated()))
    }
    ```

All these approaches are reaosnably fast, so it's unlikely they'll have any impact on performance unless called in a very tight loop. 

```{r}
bench::mark(
  version = tidyr_new_interface(),
  exists =  tidyr_new_interface1(),
  formals = tidyr_new_interface2() 
)[1:5]
```

If you need to use `packageVersion()` inside a performance sensitive function, I recommend caching the result in `.onLoad()` (which, by convention, lives in `zzz.R`). There a few ways to do this; the following block shows one approach that matches the interface of the other ways:

```{r}
tidyr_new_interface <- function() FALSE
.onLoad <- function(...) {
  if (utils::packageVersion("tidyr") > "0.8.2") {
    tidyr_new_interface <<- function() TRUE
  }
}
```

## Testing with multiple package versions

It's good practice to test both old and new versions of the code, although you won't be able to ever run both sets of tests in the same R session. The easiest way to make sure that both versions are working and stay working is to use Travis. Here I assume you're already using Travis.

Before the package is released, you can manually install the development version using `remotes::install_github()`:

```yaml
matrix:
  include:
  - r: release
    name: tidyr-devel
    before_script: Rscript -e "remotes::install_github('tidyverse/tidyr')"
```

Once the package has been released, you can remove this build. 

It's not generally that important to check that your code continues to work with an older version of the package, but if you want to you can use `remotes::install_version()`:

```yaml
matrix:
  include:
  - r: release
    name: tidyr-0.8
    before_script: Rscript -e "remotes::install_version('tidyr', '0.8.3')"
```

## Removing the conditions

Once a sufficient amount of time has elapsed (a year after release?), you should bump the required version of the package in your DESCRIPTION and remove all the branches from the code. This is easiest if you've used a single version; just search for it in the source.