# Work with multiple dependency versions {#changes-multivers}

```{r, include = FALSE}
source("common.R")
```

## What's the pattern?

In an ideal world, when a dependency of your package changes its interface, you want your package to work with both versions. This is more work but it has two significant advantages:

* The CRAN submission process is decoupled. If your package only works with the
  development version of a dependency, you'll need to carefully coordinate your
  CRAN submission with the dependencies CRAN submission. If your package works
  with both versions, you can submit first, making life easier for CRAN and for
  the maintainer of the dependency.
  
* User code is less likely to be effected. If your package only works with the
  latest version of the dependency, then when a user upgrades your package, the
  dependency also must update. Upgrading multiple packages is more likely to
  affect user code than updating a single package.

In this pattern, you'll learn how to write code designed to work with multiple versions of a dependency, and you'll how to adapt your existing Travis configuration to test that you've got it right.

## Writing code

Sometimes there will be an easy way to change the code to work with both old and new versions of the package. However, in most cases, you'll need an `if` statement that uses one call for old versions of the dependency and one call for new versions of the package:

```{r, eval = FALSE}
if (dependency_has_new_interface()) {
  # freshly written code that works with new version
} else {
  # previously written code
}
```

There are three basics approaches to test if a package has the new interface:

* Check for existence of a function.
* Check for existence or value of an argument.
* Check the version of the package.

Checking the version of the package has significant downsides, so I don't recommend it. We'll come back to the problems shortly. 

To make the problem concrete so we can show of some real code, lets imagine we have a package that uses `tidyr::nest()`. Between versions 0.8.3 and 1.0.0, tidyr changed the interface of `nest()`, so packages that use these functions need to include code like this:

```{r, eval = FALSE}
if (tidyr_new_interface()) {
  out <- tidyr::nest_legacy(df, x, y, z)
} else {
  out <- tidyr::nest(df, c(x, y, z))
}
```

Note that this code will generate an an R CMD check `NOTE` when you submit it to CRAN because `unnest_legacy()` will not be found in the CRAN version of tidyr. This is one of the few NOTEs that you can explain to CRAN: just mention that it is needed for forward compatibility in your submission notes.

The main question is how to write `tidyr_new_interface()`. There are two options:

*   As you can see above, the new version of tidyr provides `unnest_legacy()`  
    so that if you want a quick fix you can just change the function name. This
    means that we could check for the existence of that function:
  
    ```{r}
    tidyr_new_interface <- function() {
      exists("unnest_legacy", asNamespace("tidyr"))
    }
    ```
  
*   Another approach that's useful when a single argument is deprecated is to
    check for the sentinel value that's used to advertise that the argument is
    deprecated. If the package uses [lifecycle](http://lifecycle.r-lib.org/)
    then you can test for the presence of `deprecated()`:
    
    ```{r}
    tidyr_new_interface2 <- function() {
      identical(formals(tidyr::nest)$.key, quote(deprecated()))
    }
    ```

Both approaches are fast, and it's unlikely they'll have any impact on performance unless called in a very tight loop. 

```{r}
bench::mark(
  exists =  tidyr_new_interface(),
  formals = tidyr_new_interface2() 
)[1:5]
```

### Why not test the package version?

You might wonder why we don't recommend the use of `packageVersion()`, as it seems like a straightforward solution. The first problem is that it's quite slow:

```{r}
bench::mark(
  version = utils::packageVersion("tidyr") > "0.8.99"
)[1:5]
```

This means that you have to use it with with care if it's called inside a loop or in a performance sensitive function. If you do want to use `packageVersion()` I recommend caching the result `.onLoad()` (which, by convention, lives in `zzz.R`). There a few ways to do this; the following block shows one approach that matches the interface of the other ways.

```{r}
tidyr_new_interface <- function() FALSE
.onLoad <- function(...) {
  if (utils::packageVersion("tidyr") > "0.8.2") {
    tidyr_new_interface <<- function() TRUE
  }
}
```

It has two main drawbacks: it's hard to get the version specification exactly right, and its relatively slow.

* `packageVersion("tidyr") >= "1.0.0"` won't work until the development version 
   is released, so you can't easily test your code, and it doesn't help with
   coupling on CRAN.
   
* `packageVersion("tidyr") > "0.8.3"` is safer, but it will fail if there's a 
  long release process, and during release we need to do a patch release.
  This doesn't happen with most packages but has happened for packages like 
  ggplot2 and dplyr which have a large number of reverse dependencies.
  
* `packageVersion("tidyr") > "0.8.99"` is safe, but not evocative.

Additionally, `packageVersion()` is quite slow:

## Testing with multiple package versions

It's good practice to test both old and new versions of the code, although you won't be able to ever run both sets of tests in the same R session. The easiest way to make sure that both versions are working and stay working is to use Travis.

Before the package is released, you can manually install the development version using `remotes::install_github()`:

```yaml
matrix:
  include:
  - r: release
    name: tidyr-devel
    before_script: Rscript -e "remotes::install_github('tidyverse/tidyr')"
```

Once the package has been released, you can remove this build. 

It's generally not worthwhile to check that your code continues to work with an older version of the package (because generally R users are effectively forced to have the latest version of packages), but if you want to you can use `remotes::install_version()`:

```yaml
matrix:
  include:
  - r: release
    name: tidyr-0.8
    before_script: Rscript -e "remotes::install_version('tidyr', '0.8.3')"
```

## Removing the branches

Once a sufficient amount of time has elapsed (a year after release?), you should bump the required version of the package in your DESCRIPTION and remove all the branches from the code. This is easiest if you've used a single version; just search for it in the source.