# Case study: `rep()` {#cs-rep}

```{r, include = FALSE}
source("common.R")
```

## What does `rep()` do?

`rep()` is an extremely useful base R function that repeats a vector `x` in various ways. It has three details arguments: `times`, `each`, and `length.out` that interact in complicated ways. Let's explore the basics first:

```{r}
x <- c(1, 2, 4)

rep(x, times = 3)
rep(x, length.out =  10)

rep(x, times = 3, length.out =  10)
```

`times` and `length.out` both replicate the vector in the same way, but `length.out` allows you to specify a non-integer number of replications. If you specify both, `length.out` wins.

The `each` argument repeats individual components of the vector rather than the whole vector:

```{r}
rep(x, each = 3)
```

And you can combine that with `times`:

```{r}
rep(x, each = 3, times = 2)
```

If you supply a vector to `times` it works a similar way to `each`, repeating each component the specified number of times:

```{r}
rep(x, times = x)
```

## What are the problems?

* `times` with vector is misleading as it behaves more like vectorised `each`.
* complicated dependency between `times`, `length.out`, and `each`.

## How might we resolve?

Rather than trying to do everything in one function with complicated dependencies between the arguments, I think there are two underlying functions:

* `rep_each(x, times)` which replicates each `times` times element.
* `rep_full(x, times)` which replicates the entire vector `times` times.

```{r}
library(vctrs)

rep_each <- function(x, times) {
  vec_assert(times, numeric())
  times <- vec_recycle(times, vec_size(x))
  
  rep.int(x, times)
}

rep_full <- function(x, times, length) {
  if (!xor(missing(times), missing(length))) {
    stop("Must supply exactly one of `times` and `length`", call. = FALSE)
  } else if (!missing(times)) {
    vec_assert(times, numeric(), 1L)
    length <- times * base::length(x)
  } else if (!missing(length)) {
    vec_assert(length, numeric(), 1L)
  }
  
  rep_len(x, length)
}
```

(note the downside of using `length` as the argument name: we have to call `base::length()` to avoid evaluating the missing `length` when times is supplied.)

```{r}
rep_each(x, 2)
rep_each(x, x)
rep_full(x, 2)
```

If you want to both replicate each component _and_ the entire vector, you use two function calls. This is obviously more verbose than `rep()`, but I don't think it's a terribly common use case. However, one argument for a single rep function is that `rep_each()` and `rep_full()` return the same result if you change their order (i.e. they're commutative):

```{r}
rep_full(rep_each(x, 2), 3)
rep_each(rep_full(x, 3), 2) 
```

The implementation of `rep_full()` and `rep_each()` in terms of `rep.int()` and `rep_len()` suggests that R-core members are aware of the problem.