# Data, dots, details {#dots-lhs}

When you use `...` in a function where should you put it? The "data, dots, details" pattern tells you: place `...` in between data and `...` arguments. This placement has two advantages:

* It forces the user of your function to fully name the detail arguments, 
  because arguments that come after `...` after never matched by position
  or by partial name. Using full names for details arguments is good practice.
  
* When coupled with "inspect the dots" (Chapter \@ref(inspect-the-dots)), it
  minimises the chances that misspelled arguments names will silently go astray.

## Case study

```{r, eval = FALSE}
median(x, na.rm = FALSE, ...)
```

Both of these are suboptimal, because if you accidentally supply a sequence of values rather than a wrapper value, (as you might expect from `sum()` and `prod()`.), the named arguments will silently swallow the extra values. This is probably not a general rule, but I think it helps to force people to fully spell out the detail argument names. Bigger problem is mismatch of API design between `

```{r}
args(sum)
args(mean)
args(getS3method("mean", "default"))
args(weighted.mean)
args(getS3method("weighted.mean", "default"))
```
