# Required args shouldn't have defaults {#def-required}

```{r, include = FALSE}
source("common.R")
```

## What's the problem?

The absence of a default value should imply than an argument is required; the presence of a default should imply that an argument is optional.

When reading a function, it's important to be able to tell at a glance which arguments must be supplied and which are optional. Otherwise you need to rely on the user having carefully read the documentation.

## What are some examples?

*   In `sample()` neither `x` not `size` has a default value, suggesting that
    both are required, and the function would error if you didn't supply them.
    But `size` is optional, determined by a complex conditional.
    
    ```{r}
    sample(1:4)
    sample(4)
    ```

*   `rt()` (draw random numbers from the t-distribution) looks like it 
    requires the `ncp` parameter but it doesn't.

*   `download.file()` looks like it requires the `method` argument but 
    actually consults a global option (`download.file.method`) if it's not
    supplied.

*   `lm()` does not have defaults for `formula`, `data`, `subset`, `weights`, 
    `na.action`, or `offset`. Only `formula` is actually required, but even 
    its absence fails to generate a clear error message:
    
    ```{r, error = TRUE}
    lm()
    ```
    
*   `help()` and `vignette()` have no default for their first argument, 
    suggesting that they're required. But they're not: calling `help()` or
    `vignette()` without any arguments lists all help topics and vignettes 
    respectively.
 
*   In `diag()`, the argument `x` has a default `1`, but it's required: if 
    you don't supply it you get an error:
    
    ```{r, error = TRUE}
    diag()
    diag(x = 1)
    ```
    
    Conversely, `nrow` and `ncol` don't have defaults but aren't required.

*   In `ggplot2::geom_abline()`, `slope` and `intercept` don't have defaults 
    but are not required. If you don't supply them they default to `slope = 1` 
    and `intercept = 0`, *or* are taken from `aes()` if they're provided
    there.

A common warning sign is the use of `missing()` inside the function.

## What are the exceptions?

There are two exceptions to this rule:

* A pair of arguments that provide an alternative specification for the same
  underlying concept. It is only ever possible to supply one argument.
  
* When you can either supply one complex object, or a handful of simpler 
  objects.

In both cases, I believe the benefits outweigh the costs of violating a standard pattern.

### Pair of mututally exclusive arguments {#args-mutually-exclusive}

A number of functions that allow you to supply exactly one of two possible arguments:

* `read.table()` allows you to supply data either with a path to a `file`,
  or inline as `text`.

* `rvest::html_node()` allows you to select HTML nodes either with a `css` 
  selector or an `xpath` expression.
  
* `forcats::fct_other()` allows you to either `keep` or `drop` specified 
  factor values.

* `modelr::seq_range()` allows you create a sequence over the range of `x` by 
  either specifying the length of the sequence (with `n`) or the distance 
  between values (with `by`).
  
If you use this technique, use `xor()` and `missing()` to check that exactly one argument is supplied:

```{r, eval = FALSE}
if (!xor(missing(keep), missing(drop))) {
  stop("Must supply exactly one of `keep` and `drop`", call. = FALSE)
}
```

And in the documentation, make it clear that only one of the pair can be supplied:

```{r}
#' @param keep,drop Pick one of `keep` and `drop`:
#'   * `keep` will preserve listed levels, replacing all others with 
#'     `other_level`.
#'   * `drop` will replace listed levels with `other_level`, keeping all
#'     as is.
```

This technique should only be used for are exactly two possible arguments. If there are more than two , that is generally a sign you should create more functions. See case studies in Chapter \@ref(cs-rep) and Section \@ref(cs-fct-lump) for examples.

### One compound argument vs multiple simple arguments {#args-compound}

A related, if less generally useful, form is to allow the user to supply either a single complex argument or several smaller arguments. For example:

* `stringr::str_sub(x, cbind(start, end))` is equivalent to 
  `str_sub(x, start, end)`.
  
* `stringr::str_replace_all(x, c(pattern = replacement))` is equivalent to
  `stringr(x, pattern, replacement)`.
  
* `rgb(cbind(r, g, b))` is equivalent to `rgb(r, g, b)` (See Chapter 
  \@ref(cs-rgb) for more details).
  
* `options(list(a = 1, b = 2))` is equivalent to `options(a = 1, b = 2)`.

The most compelling reason to provide this sort of interface is when another function might return a complex output that you want to use as an input. For example, it seems reasonable that you should be able to feed the output of `str_locate()` directly into `str_sub()`:

```{r}
library(stringr)

x <- c("aaaaab", "aaab", "ccccb")
loc <- str_locate(x, "a+b")

str_sub(x, loc)
```

But equally, it would be weird to have to provide a matrix when subsetting with known positions:

```{r}
str_sub("Hadley", cbind(2, 4))
```

So `str_sub()` allows either individual vectors supplied to `start` and `end`, or a two-colummn matrix supplied to `start`.

To implement in your own functions, you should branch on the type of the first argument:

(Why? Why not branch if the other arguments are missing? Or some combination?)

```{r}
str_sub <- function(string, start = 1L, end = -1L) {
  if (is.matrix(start)) {
    if (!missing(end)) {
      stop("`end` must be missing when `start` is a matrix", call. = FALSE)
    }
    if (ncol(start) != 2) {
      stop("Matrix `start` must have exactly two columns", call. = FALSE)
    }
    stri_sub(string, from = start[, 1], to = start[, 2])
  } else {
    stri_sub(string, from = start, to = end)
  }
}
```

And make it clear in the documentation:

```{r}
#' @param start,end Integer vectors giving the `start` (default: first)
#'   and `end` (default: last) positions, inclusively. Alternatively, you
#'   pass a two-column matrix to `start`, i.e. `str_sub(x, start, end)`
#'   is equivalent to `str_sub(x, cbind(start, end))`
```
