# Explain important defaults {#def-inform}

```{r, include = FALSE}
source("common.R")
library(dplyr, warn.conflicts = FALSE)
```

## What's the pattern?

If a default value is important, and the computation is non-trivial, inform the user what value was used. This is particularly important when the default value is an educated guess, and you want the user to change it.

## What are some examples?

*   `dplyr::left_join()` and friends automatically compute the variables to 
    join `by` as the variables that occur in both `x` and `y` (this is called 
    a natural join in SQL). This is convenient, but it's a heuristic so 
    doesn't always work.
    
    ```{r, error = TRUE}
    library(nycflights13)
    library(dplyr)
    
    # Correct    
    out <- left_join(flights, airlines)
    
    # Incorrect
    out <- left_join(flights, planes)
    
    # Error
    out <- left_join(flights, airports)
    ```

*   `readr::read_csv()` reads a csv file into a data frame. Because csv files 
    don't store the type of each variable, readr must guess the types. In order 
    to be fast, `read_csv()` uses some heuristics, so it might guess wrong. Or 
    maybe guesses correctly today, but when your automated script runs in two 
    months time when the data format has changed, it might guess incorrectly
    and give weird downstream errors. For this reason, `read_csv()` prints
    the column specification in a way that you can copy-and-paste into your 
    code.

    ```{r}
    library(readr)
    mtcars <- read_csv(readr_example("mtcars.csv"))
    ```

*   In `ggplot2::geom_histogram()`, the `binwidth` is an important parameter 
    that you should always experiment with. This suggests it should be a 
    required argument, but it's hard to know what values to try until you've 
    seen a plot. For this reason, ggplot2 provides a suboptimal default of
    30 bins: this gets you started, and then a message tells you how to modify.
  
    ```{r}
    library(ggplot2)
    ggplot(diamonds, aes(carat)) + geom_histogram()
    ```

*  When installing packages, `install.packages()` informs of the value of the
   `lib` argument, which defaults to `.libPath()[[1]]`:

    ```{r, eval = FALSE}
    install.packages("forcats")
    # Installing package into ‘/Users/hadley/R’
    # (as ‘lib’ is unspecified)
    ```
    
    This, however, is not terribly important (most people only use one library),
    it's easy to ignore this amongst the other output, and the message doesn't 
    refer to the mechanism that controls the default (`.libPaths()`).

## Why is it important?

This is particulary important for descriptor arguments (Chapter \@ref(args-data-details)). 

Need to balance with the counterpoint that people don't read output that is constantly repeated. (For example, do you know how to cite R in a paper? It's mentioned every time that you start R.)

## How can I use it?

There are two main principles to adhere to when use this technique:

* Be silenced when a default value is provided.

* Include code to copy and paste to avoid it in the future.

Generally, you'll implement these through a helper function. It should compute the default value, and generating a `message()` displaying the result. Take the dplyr join functions, for example. They use a function like this `common_by()`:

```{r}
common_by <- function(x, y) {
  common <- intersect(names(x), names(y))
  if (length(common) == 0) {
    stop("Must specify `by` when no common variables in `x` and `y`", call. = FALSE)
  }
  
  message("Computing common variables: `by = ", rlang::expr_text(common), "`")
  common
}

common_by(data.frame(x = 1), data.frame(x = 1))
common_by(flights, planes)
```

The technique you use to generate the code will vary from function to function. `rlang::expr_text()` is useful here because it automatically creates the code you'd use to build the character vector.

You can use the function directly as the default value, `by = common_by(x, y)`, which is most appropriate if you want to export and document it. Otherwise, to avoid creating a magical default (Chapter \@ref(def-magical)), use the technique of Section \@ref(arg-short-null):

```{r}
left_join <- function(x, y, by = NULL) {
  by <- by %||% common_by(x, y)
}
```

If you use this function in multiple places, it may be helpful to parameterise the argument name:

```{r}
common_by <- function(x, y, arg = "by") {
  common <- intersect(names(x), names(y))
  if (length(common) == 0) {
    stop(
      glue::glue("Must specify `{arg}` when no common variables in `x` and `y`"),
      call. = FALSE
    )
  }
  
  message(glue::glue("Computing common variables: `{arg} = {deparse(common)}`"))
  common
}

common_by(flights, planes)
```
