# Dots

```{r, include = FALSE}
source("common.R")
```

* Allow S3 methods to take additional arguments.
* Pass on to another function (particularly in functional programming).
* Create data structures.

### Avoiding false matches

Place as far to the left as possible to force users to name arguments. Should come between data and details arguments.

Use `ellipsis::check_dots_unnamed()` or `ellipsis::check_dots_used()`

Ensure arguments to the right of the dots (i.e. details arguments) are prefixed with `.`. (The dots prefix is not considered part of the argument name so it's ok to have some functions with argument `f` and some with `.f`.)

### Don't use just to avoid `c()` {#vectors-not-dots}

i.e. `forcats::fct_relevel()` as it makes future API changes quite difficult. 

### Allow trailing dot

Use `list2()` instead of list.

```{r, error = TRUE}
list(1, 2, )
rlang::list2(1, 2, )
```

If you want something simpler of guaranteed type, you have to use `vctrs::vec_c(.ptype = integer())`. If you don't want a vctrs dependency, you'll need `c()` + an error if it's the incorrect type.

Other rlang capturing functions 
