# Put `…` after required arguments {#sec-dots-after-required}

```{r}
#| include = FALSE
source("common.R")
```

## What's the pattern?

If you use `…` in a function, put it after the required arguments and before the optional arguments.

This has two positive impacts:

-   It forces the user of your function to fully name optional arguments, because arguments that come after `...` are never matched by position or by partial name.
    We believe that using full names for optional arguments is good practice because it makes code easier to read.

-   This in turn means that uou can easily add new optional arguments or change the order of existing arguments without affecting existing code.

## What are some examples?

The arguments to `mean()` are `x`, `trim`, `na.rm` and `…`.
This means that you can write code like this:

```{r}
x <- c(1, 2, 10, NA)
mean(x, , TRUE)
mean(x, n = TRUE, t = 0.1)
```

Not only does this allow for confusing code[^dots-after-required-1], it also makes it hard to later change the order of these arguments, or introduce new arguments that might be more important.

[^dots-after-required-1]: As much as we recommended people don't write code like this, you know someone will!

If `mean()` instead placed `…` before `trim` and `na.rm`, like `mean2()`[^dots-after-required-2] below, then you must fully name each argument:

[^dots-after-required-2]: Note that I moved `na.rm = TRUE` in front of `trim` because I believe `na.rm` is the more important argument because it's used vastly more often than `trim` and I'm following @sec-important-args-first.

```{r}
mean2 <- function(x, ..., na.rm = FALSE, trim = 0) {
  mean(x, ..., na.rm = na.rm, trim = trim)
}

mean2(x, na.rm = TRUE)
mean2(x, na.rm = TRUE, trim = 0.1)
```

## How do I remediate past mistakes?

It's straightforward to fix a function where you've put `...` in the wrong place: you just need to change the argument order and use `rlang::check_dots_used()` to check that no arguments are lost (learn more in @sec-dots-inspect).
This is a breaking change, but it tends to affect relatively little code because most people do fully name optional arguments.

We can use this approach to make a safer version of `mean()`:

```{r}
#| error = TRUE
mean3 <- function(x, ..., na.rm = FALSE, trim = 0) {
  rlang::check_dots_used()
  mean(x, ..., na.rm = na.rm, trim = trim)
}

mean3(x, , TRUE)

mean3(x, n = TRUE, t = 0.1)
```

::: {.callout-note collapse="true"}
## Base R

In base R you can use `base::chkDots()`, but it uses a slightly simpler technique which means it's not suitable for usage in S3 methods.
:::

## See also

-   @sec-dots-data: if `…` is a required argument because it's used to combine an arbitrary number of objects in a data structure.
-   @sec-dots-inspect: to ensure that arguments to `…` never go silently missing.
