# Put `…` before optional arguments {#sec-dots-position}

```{r}
#| include = FALSE
source("common.R")
```

## What's the pattern?

If you use `…` in a function, put it after the required arguments and before the optional arguments.

(See @sec-dots-data if `…` requires at least one argument because it's used to combine an arbitrary number of objects.)

## Why is it important?

There are three primary advantages:

-   It forces the user of your function to fully name optional arguments, because arguments that come after `...` are never matched by position or partially by name.
    Using full names for details arguments is good practice, because we believe it makes code easier to read.

-   You can easily add new optional arguments or change the order of existing arguments without affecting existing calls.
    This makes it easy to extend your function with new capabilities, because you don't need to worry about breaking existing code.

-   When coupled with "inspect the dots" (@sec-dots-inspect), or "dot prefix" (@sec-dots-prefix) it minimises the chances that misspelled arguments names will silently go astray.

The payoff of this pattern is not huge: it protects against a fairly unusual failure mode.
However, since the failure mode is silent (and hence easy to miss) and the pattern is very easy to apply, I think it's still worth it.

## What are some examples?

The arguments to `mean()` are `x`, `trim`, `na.rm` and `…`.
This means that you can write code like this:

```{r}
x <- c(1, 2, 10, NA)
mean(x, , TRUE)
mean(x, n = TRUE, t = 0.1)
```

Not only does allow confusing code[^dots-position-1], it also makes it hard to later change the order of these arguments, or introduce new arguments that might be more important.

[^dots-position-1]: As much as we recommended people don't write code like this, you know someone will!

If `mean()` instead placed `…` before `trim` and `na.rm`, like `mean2()`[^dots-position-2] below, then you must fully name each argument:

[^dots-position-2]: Note that I moved `na.rm = TRUE` in front of `trim` because I believe `na.rm` is the more important argument because it's used vastly more often than `trim` and I'm following @sec-args-data-details.

```{r}
mean2 <- function(x, ..., na.rm = FALSE, trim = 0) {
  mean(x, ..., na.rm = na.rm, trim = trim)
}

mean2(x, na.rm = TRUE)
mean2(x, na.rm = TRUE, trim = 0.1)
```

## How do I remediate past mistakes?

If you've already published a function where you've put `...` in the wrong place, it's easy to fix.
You'll need to use a function from the rlang package to check that `...` is as expected (e.g. from @sec-dots-data or @sec-dots-inspect).
Since using the full names for details arguments is good practice, making this change will typically affect little existing code, but it is an interface change so should be advertised prominently.

```{r}
#| error = TRUE

old_interface <- function(x, data1 = 1, data2 = 2, ...) {
}
old_interface(1, 2)

new_interface <- function(x, ..., data1 = 1, data2 = 2) {
  rlang::check_dots_used()
}
new_interface(1, 2)
```

We can use this approach to make a safer version of `mean()`:

```{r}
#| error = TRUE
mean3 <- function(x, ..., na.rm = FALSE, trim = 0) {
  rlang::check_dots_used()
  mean(x, ..., na.rm = na.rm, trim = trim)
}

mean3(x, , TRUE)

mean3(x, n = TRUE, t = 0.1)
```
