# Error constructors {#err-constructor}

```{r, include = FALSE}
source("common.R")
```

## What's the pattern?

Following the rule of three, whenever you generate the same error in three or more places, you should extract it out into a common function, called an __error constructor__. This function should create a [custom condition](https://adv-r.hadley.nz/conditions.html#custom-conditions) that contains components that can easily be tested and a `conditionMessage()` method that generates user friendly error messages.

(This is a new pattern that we are currently rolling out across the tidyverse; it's currently found in few packages.)

```{r setup}
library(rlang)
```

## Why is this important?

* If you don't use an custom condition, you can only check that your function
  has generated the correct error by matching the text of the error message
  with a regular expression. This is fragile because the text of error messages
  changes relatively frequently, causing spurious test failures.

* You _can_ use custom conditions for one-off errors, but generally the extra 
  implementation work is not worth the pay off. That's why we recommend only 
  using an error constructor for repeated errors.

* It gives more precise control over error handling with `tryCatch()`. 
  This is particularly useful in packages because you may be able to give
  more useful high-level error mesasges by wrapping a specific low-level 
  error. 

* As you start using this technique for more error messages you can create
  a hierarchy of errors that allows you to borrow behaviour, reducing the
  amount of code you need to write.

* Once you have identified all the errors that can be thrown by a function,
  you can add a `@section Throws:` to the documentation that precisely 
  describes the possible failure modes.

## What does an error constructor do?

An error constructor is very similar to an [S3 constructor](https://adv-r.hadley.nz/s3.html#s3-constructor), as its job is to extract out repeated code and generate a rich object that can easily be computed with. The primary difference is that instead of creating and returning a new object, it creates a custom error and immediately throws it with `abort()`.

Here's a simple imaginary error that might be thrown by [fs](http://fs.r-lib.org/) if it couldn't find a file:

```{r}
stop_not_found <- function(path) {
  abort(
    .subclass = "fs_error_not_found",
    path = path
  )
}
```

Note the naming scheme:

* The function should be called `stop_{error_type}`

* The error class should be `{package}_error_{error_type}`. 

The function should have one argument for each varying part of the error, and these argument should be passed onto `abort()` to be stored in the condition object.

To generate the error message shown to the user, provide a `conditionMessage()` method:

```{r}
#' @export
conditionMessage.fs_error_not_found <- function(cnd) {
  glue::glue_data(cnd, "'{path}' not found")
}
```

```{r, include = FALSE}
vctrs::s3_register("base::conditionMessage", "fs_error_not_found")
```


```{r, eval = FALSE}
stop_not_found("a.csv")
#> Error: 'a.csv' not found
```

This method must be exported, because you are defining a method for a generic in another package, and it will often use `glue::glue_data()` to assemble the components of the condition into a string. See <https://style.tidyverse.org/error-messages.html> for advice on writing the error message.

## How do I test?

```{r}
library(testthat)
```


### Test the constructor

Firstly, you should test the error constructor. The primary goal of this test is to ensure that the error constructor generates a message that is useful to humans, which you can not automate. This means that you can not use a unit test (because the desired output is not known) and instead you need to use a regression test, so you can ensure that the message does not change unexpectedly. For that reason the best approach is usually to use [`verify_output()`](https://testthat.r-lib.org/reference/verify_output.html), e.g.:

```{r, eval = FALSE}
test_that("stop_not_found() generates useful error message", {
  verify_output(test_path("test-stop-not-found.txt"), {
    stop_not_found("a.csv")
  })
})
```

This is useful for pull requests because `verify_output()` generates a complete error messages in a text file that can easily be read and reviewed.

If your error has multiple arguments, or your `conditionMessage()` method contains `if` statements, you should generally attempt to cover them all in a test case. 

### Test usage

Now that you have an error constructor, you'll need to slightly change how you test your functions that use the error constructor. For example, take this imaginary example for reading a file into a single string:

```{r}
read_lines <- function(x) {
  if (!file.exists(x)) {
    stop_not_found(x)
  }
  paste0(readLines(x), collapse = "\n")
}
```

Previously, you might have written:

```{r}
expect_error(read_lines("missing-file.txt"), "not found")
```

But, now as you see, testthat gives you a warning that suggests you need to use the class argument instead:

```{r}
expect_error(read_lines("missing-file.txt"), class = "fs_error_not_found")
```

This is less fragile because you can now change the error message without having to worry about breaking existing tests.

If you also want to check components of the error object, note that `expect_error()` returns it:

```{r}
cnd <- expect_error(read_lines("missing-file.txt"), class = "fs_error_not_found")
expect_equal(cnd$path, "missing-file.txt")
```

I don't think this level of testing is generally important, so you should only use it because the error generation code is complex conditions, or you have identified a bug.

## Error hierarchies

As you start writing more and more error constructors, you may notice that you are starting to share code between them because the errors form a natural hierarchy. To take advantage of this hierarchy to reduce the amount of code you need to write, you can make the errors subclassable by adding `...` and `class` arguments:

```{r}
stop_not_found <- function(path, ..., class = character()) {
  abort(
    .subclass = c(class, "fs_error_not_found"),
    path = path
  )
}
```

Then the subclasses can call this constructor, and the problem becomes one of S3 class design. We currently have little experience with this, so use with caution.