# Argument meaning should be independent {#sec-independent-meaning}

```{r}
#| include = FALSE
source("common.R")
```

## What's the problem?

Avoid having one argument change the interpretation of another argument.

This makes reading function calls harder because if the argument that changes meaning comes after the argument who's meaning is changed you need to reinterpret the rest of the call.

## What are some examples?

-   In `library()` the `character.only` argument changes how the `package` argument is interpreted:

    ```{r}
    #| eval = FALSE
    ggplot2 <- "dplyr"

    # Loads ggplot2
    library(ggplot2)

    # Loads dplyr
    library(ggplot2, character.only = TRUE)
    ```

-   In `install.packages()` setting `repos = NULL` changes the interpretation of `pkgs` from being a vector of package names to a vector of file paths.

-   In the base R string functions, `perl` and `fixed` change the interpretation of the `pattern` argument.
    See @sec-pattern-engine for more details.

-   In `ggplot2::geom_label()`, setting `parse = TRUE` changes the meaning of the label argument/aesthetic from being any string to be a string on unparsed R code.

-   In `readr::locale()` there's a complex dependency between `decimal_mark` and `grouping_mark` because they can't be the same value, and Europe and the US Europe use different standards.
    That suggests it might have been better to specify them in a single argument like `marks = c(",", ".")`.

-   In `findInterval()` if you set `left.open = TRUE` the `rightmost.closed` means `leftmost.closed`.

## How do I remediate past mistakes?

## See also

@sec-independent-usage discusses the related problem of when using one argument affects which other arguments can be used.
