# Mutually exclusive arguments {#sec-mutually-exclusive}

```{r}
#| include = FALSE
source("common.R")
```

## What's the pattern?

## What are some examples?

A number of functions that allow you to supply exactly one of two possible arguments:

-   `read.table()` allows you to supply data either with a path to a `file`, or inline as `text`.

-   `rvest::html_element()` allows you to select HTML elements either with a `css` selector or an `xpath` expression.

-   In `ggplot2::scale_x_date()` and friends you can specify the breaks and labels either with `breaks` and `labels` (like all other scale functions) or with `date_breaks` and `date_labels`.
    If you set both values in a pair, the `date_` version wins.

-   `forcats::fct_other()` allows you to either `keep` or `drop` specified factor values.

-   In `var()` you can remove missing values by either setting `na.rm = TRUE` or `use = "complete.obs"`.
    If you set them both, `use` wins.

## How do you use this pattern?

If you use this technique, use `xor()` and `missing()` to check that exactly one argument is supplied:

```{r}
#| error: true

fct_drop <- function(f, drop, keep) {
  if (!xor(missing(keep), missing(drop))) {
    stop("Must supply exactly one of `keep` and `drop`")
  }  
}
fct_drop(factor())

fct_drop(factor(), keep = "a", drop = "b")
```

Or the helper `rlang::check_exclusive()`:

```{r}
#| error: true

fct_drop <- function(f, drop, keep) {
  rlang::check_exclusive(drop, keep)
}
fct_drop(factor())

fct_drop(factor(), keep = "a", drop = "b")
```

And in the documentation, make it clear that only one of the pair can be supplied:

```{r}
#' @param keep,drop Pick one of `keep` and `drop`:
#'   * `keep` will preserve listed levels, replacing all others with 
#'     `other_level`.
#'   * `drop` will replace listed levels with `other_level`, keeping all
#'     as is.
```

This technique should only be used for are exactly two possible arguments.
If there are more than two, that is generally a sign you should create more functions.
See case studies in @sec-cs-rep and @sec-cs-fct-lump for examples.
