# Type-stability {#out-type-stability}

```{r, include = FALSE}
source("common.R")
```

The less you need to know about a function's inputs to predict the type of its output, the better. Ideally, a function should either always return the same type of thing, or return something that can be trivially computed from its inputs.

If a function is __type-stable__ it satisfies two conditions:

* You can predict the output type based only on the input types 
  (not their values).
  
* If the function uses `...`, the order of arguments in does not affect the 
  output type.

```{r setup}
library(vctrs)
```

## Simple examples

*   `purrr::map()` and `base::lapply()` are trivially type-stable because they
    always return lists.

*   `paste()` is type stable because it always returns a character vector.

    ```{r}
    vec_ptype(paste(1))
    vec_ptype(paste("x"))
    ```

*   `base::mean(x)` almost always returns the same type of output as `x`. 
    For example, the mean of a numeric vector is a numeric vector, and the mean
    of a date-time is a date-time.
    
    ```{r}
    vec_ptype(mean(1))
    vec_ptype(mean(Sys.time()))
    ```

*   `ifelse()` is not type-stable because the output type depends on the value:

    ```{r}
    vec_ptype(ifelse(NA, 1L, 2))
    vec_ptype(ifelse(FALSE, 1L, 2))
    vec_ptype(ifelse(TRUE, 1L, 2))
    ```

## More complicated examples

Some functions are more complex because they take multiple input types and have to return a single output type. This includes functions like `c()` and `ifelse()`. The rules governing base R functions are idiosyncratic, and each function tends to apply it's own slightly different set of rules. Tidy functions should use the consistent set of rules provided by the [vctrs](https://vctrs.r-lib.org) package.

## Challenge: the median

A more challenging example is `median()`. The median of a vector is a value that (as evenly as possible) splits the vector into a lower half and an upper half. In the absence of ties, `mean(x > median(x)) == mean(x <= median(x)) == 0.5`. The median is straightforward to compute for odd lengths: you simply order the vector and pick the value in the middle, i.e. `sort(x)[(length(x) - 1) / 2]`. It's clear that the type of the output should be the same type as `x`, and this algorithm can be applied to any vector that can be ordered.

But what if the vector has an even length? In this case, there's no longer a unique median, and by convention we usually take the mean of the middle two numbers. 

In R, this makes the `median()` not type-stable:

```{r}
typeof(median(1:3))
typeof(median(1:4))
```

Base R doesn't appear to follow a consistent principle when computing the median of a vector of length 2. Factors throw an error, but dates do not (even though there's no date half way between two days that differ by an odd number of days).

```{r, error = TRUE}
median(factor(1:2))
median(Sys.Date() + 0:1)
```

To be clear, the problems caused by this behaviour are quite small in practice, but it makes the analysis of `median()` more complex, and it makes it difficult to decide what principle you should adhere to when creating `median` methods for new vector classes.

```{r, error = TRUE}
median("foo")
median(c("foo", "bar"))
```

## Exercises

1.  How is a date like an integer? Why is this inconsistent? 

    ```{r}
    vec_ptype(mean(Sys.Date()))
    vec_ptype(mean(1L))
    ```

