# Side-effect functions should return invisibly {#out-invisible}

```{r, include = FALSE}
source("common.R")
```

## What's the pattern?

If a function is called primarily for its side-effects, it should invisibly return a useful output. If there's no obvious output, return the first argument. This makes it possible to use the function with in a pipeline.

## What are some examples?

```{r, eval = FALSE, include = FALSE}
source("fun_def.R")
pkg_funs("base") %>% 
  funs_body_keep(has_call, "invisible") %>% 
  discard(~ grepl("print", .x$name))
```

* `print(x)` invisibly returns the printed object.

* `x <- y` invisible returns `y`. This is what makes it possible to chain
  together multiple assignments `x <- y <- z <- 1`

* `readr::write_csv()` invisibly returns the data frame that was saved.

* `purrr::walk()` invisibly returns the vector iterated over.

* `fs:file_copy(from, to)` returns `to`

* `options()` and `par()` invisibly return the previous value so you can 
  reset with `on.exit()`.

## Why is it important?

Invisibly returning the first argument allows to call the function mid-pipe for its side-effects while allow the primary data to continue flowing through the pipe. This is useful for generating intermediate diagnostics, or for saving multiple output formats.

```{r}
library(dplyr, warn.conflicts = FALSE)
library(tibble)

mtcars %>%
  as_tibble() %>% 
  filter(cyl == 6) %>% 
  print() %>% 
  group_by(vs) %>% 
  summarise(mpg = mean(mpg))
```

```{r}
library(readr)

mtcars %>% 
  write_csv("mtcars.csv") %>% 
  write_tsv("mtcars.tsv")

unlink(c("mtcars.csv", "mtcars.tsv"))
```

```{r}
library(fs)

paths <- file_temp() %>%
  dir_create() %>%
  path(letters[1:5]) %>%
  file_create()
paths
```

Functions that modify some global state, like `options()` or `par()`, should return the _previous_ value of the variables. This, in combination with Section \@ref(args-compound), makes it possible to easily reset the effect of the change:

```{r}
x <- runif(1)
old <- options(digits = 3)
x

options(old)
x
```
